Determination of the \ce{^{64}Zn},\ce{^{47}Ti}(n,p) Cross-Sections using a DD Neutron Generator for Medical Isotope Studies

### 2017-03-02 First_np_Paper/first_np_paper.tex
Adding citations, squashing comments

Two other potential neutron sources for (n,x) reactions exist in addition to the secondary neutron fields generated at existing isotope production facilities: reactors and neutron generators that utilize the D(T,n)$\alpha$ \enquote{DT} and D(D,n)\ce{^3He} \enquote{DD} reactions.  While reactors produce copious quantities of neutrons, their energy spectrum is often not well-suited to the preparation of high-purity samples due to the co-production of unwanted activities via neutron capture, in addition to the significant start-up costs and proliferation concerns involved in their commissioning\cite{Updegraff2013}.  Similarly, while the higher energy 14-15 MeV neutrons produced at DT generators are capable of initiating (n,p) and (n,$\alpha$) reactions, their higher energy opens the possibility of creating unwanted activities via (n,pxn) and (n,$\alpha$xn) reactions that cannot easily be separated from the desired radionuclides.

@@

The present work features a pair of cross section measurements for the production of two non-standard positron emitters: \ce{^{64}Zn}(n,p)\ce{^{64}Cu} and \ce{^{47}Ti}(n,p)\ce{^{47}Sc}. \ce{^{64}Cu}  ($t_{1/2}$ = 12.7 h) undergoes $\beta^+$ decay (61.5\% branching ratio) to \ce{^{64}Ni} or $\beta^-$ decay (38.5\% branching ratio) to \ce{^{64}Zn}. The emitted 190-keV $\beta^-$ particle makes this an  attractive short-range therapeutic radionuclide, which also has the possibility for simultaneous positron emission tomography (PET) imaging for real-time dose monitoring and verification. This makes \ce{^{64}Cu} particularly attractive for emerging radiation therapy protocols\cite{Lewis2003,NSACIsotopesSubcommittee2015,Bandari2014,mp500671j}. In addition, copper radiochemistry is well developed, and many existing ligands and carriers may be used for selective delivery of the radionuclide to different sites in patients. The second radionuclide studied, \ce{^{47}Sc} ($t_{1/2}$ = 3.35 d), undergoes $\beta^-$ decay to \ce{^{47}Ti}, emitting a high-intensity 159-keV gamma ray in the process. This radionuclide is  attractive as an emerging diagnostic isotope, due to the similarity of the emitted gamma ray to that of the more well-established \ce{^{99m}Tc}\cite{Qaim2011,Qaim201731}. However, due to the short half-life of and dwindling supplies of \ce{^{99m}Tc}, \ce{^{47}Sc} stands poised as a potential solution to this shortage, due to its longer half-life and multiple production pathways without the need for highly enriched uranium. In addition, when paired with \ce{^{44}Sc}, \ce{^{47}Sc} forms a promising \enquote{theranostic} set for use in simultaneous therapeutic and diagnostic applications.

\comment{cite theranostic papers, etc}

@@

\comment{Regex to replace all hard figure references with LaTeX cross-references}

@@

The approach used in both measurements was to irradiate foils of zinc or titanium, which were co-loaded with indium foils in order to determine their (n,p) cross sections relative to the well-established \ce{^{113}In}(n,n')\ce{^{113m}In} and \ce{^{115}In}(n,n')\ce{^{115m}In} neutron dosimetry standards.   \autoref{tab:foil_specs} records physical characteristics of each foil for the various irradiations. In each experiment, the co-loaded foils were irradiated for 3 hours at nominal operating conditions of 1.3 mA and 100 kV.  After irradiation, the foils were removed and placed in front of an appropriate High-Purity Germanium gamma-ray detector and the decay gamma-rays were recorded in variable-time bins.

@@

\comment{Regex to replace table hard link with LaTeX cross-reference.}

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
% \usepackage{graphicx}
\begin{table*}
\centering
\caption{Foil characteristics for each of the three (Zn/In)* experiments and the two (Ti/In)$^\dagger$  experiments. 
}
\label{tab:foil_specs}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}ccccccc@{}}
\toprule
Foils Used & Metal Purity          & Abundance (a/o)                                                           & \begin{tabular}[c]{@{}c@{}}Foil Density \\ (mg/cm$^2$)\end{tabular} & Thickness (mm)                                                                                                         & Diameter (cm)                                                                                                           & Mass (g)                                                                                    \\ \midrule
$^\text{nat}$In      & \textgreater 99.999\% & \begin{tabular}[c]{@{}c@{}}\ce{^{113}In} (4.29\%),\\ \ce{^{115}In} (95.71\%)\end{tabular} & 365.5                                                            & \begin{tabular}[c]{@{}c@{}}0.49 $\pm$ 0.02*, \\ 0.50 $\pm$ 0.03*, \\ 0.49 $\pm$ 0.03*, \\ 0.53 $\pm$ 0.06$^\dagger$, \\ XXX $\pm$ XXX$^\dagger$\end{tabular} & \begin{tabular}[c]{@{}c@{}}9.75 $\pm$ 0.09*, \\ 9.98 $\pm$ 0.15*, \\ 9.96 $\pm$ 0.10*, \\ 10.01 $\pm$ 0.11$^\dagger$, \\ XXX $\pm$ XXX$^\dagger$\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.248*, \\ 0.248*, \\ 0.241*, \\ 0.247$^\dagger$,\\  XXXXXX$^\dagger$\end{tabular} \\
$^\text{nat}$Zn      & \textgreater 99.99\%  & \ce{^{64}Zn} (49.17\%)                                                            & 714.1                                                            & \begin{tabular}[c]{@{}c@{}}1.03 $\pm$ 0.01, \\ 1.03 $\pm$ 0.01, \\ 1.02 $\pm$ 0.01\end{tabular}                                    & \begin{tabular}[c]{@{}c@{}}9.93 $\pm$ 0.14, \\ 9.76 $\pm$ 0.17, \\ 9.89 $\pm$ 0.15\end{tabular}                                     & \begin{tabular}[c]{@{}c@{}}0.538, \\ 0.451, \\ 0.452\end{tabular}                           \\
$^\text{nat}$Ti      & 99.999\%              & \ce{^{47}Ti} (7.44\%)                                                             & 450.6                                                            & \begin{tabular}[c]{@{}c@{}}1.16 $\pm$ 0.02, \\ XXXXXX\end{tabular}                                                         & \begin{tabular}[c]{@{}c@{}}9.93 $\pm$ 0.04, \\ XXXXXX\end{tabular}                                                          & \begin{tabular}[c]{@{}c@{}}0.337, \\ XXXXXXX\end{tabular}                                   \\ \bottomrule
\end{tabular}%
}
\end{table*}

\comment{Need foil data from Mauricio's 2nd Ti run}

@@

\comment{Regex to replace equations with LaTeX formats, add cross-references.}

@@

M. B. Chadwick,et al., Nucl. Data Sheets, vol. 107, no. 12, 2006.

@@

C. Müller et al., J. Nucl. Med., vol. 55, no. 10, Oct. 2014.

N. Otuka et al., Nucl. Data Sheets, vol. 120, 2014.

L. Pietrelli et al., J. Radioanal. Nucl. Chem., vol. 157, no. 2, 1992.

S. M. Qaim et al., IAEA Technical Reports Series No. 473, 2011.

@@

D. Updegraff et al., ''Nuclear Medicine without Nuclear Reactors or Uranium Enrichment,'' 2013.

>>>>

Two other potential neutron sources for (n,x) reactions exist in addition to the secondary neutron fields generated at existing isotope production facilities: reactors and neutron generators that utilize the D(T,n)$\alpha$ \enquote{DT} and D(D,n)\ce{^3He} \enquote{DD} reactions.  While reactors produce copious quantities of neutrons, their energy spectrum is often not well-suited to the preparation of high-purity samples due to the co-production of unwanted activities via neutron capture, in addition to the significant start-up costs and proliferation concerns involved in their commissioning \cite{Updegraff2013}.  Similarly, while the higher energy 14-15 MeV neutrons produced at DT generators are capable of initiating (n,p) and (n,$\alpha$) reactions, their higher energy opens the possibility of creating unwanted activities via (n,pxn) and (n,$\alpha$xn) reactions that cannot easily be separated from the desired radionuclides.

@@

The present work features a pair of cross section measurements for the production of two non-standard positron emitters: \ce{^{64}Zn}(n,p)\ce{^{64}Cu} and \ce{^{47}Ti}(n,p)\ce{^{47}Sc}. \ce{^{64}Cu}  ($t_{1/2}$ = 12.7 h) undergoes $\beta^+$ decay (61.5\% branching ratio) to \ce{^{64}Ni} or $\beta^-$ decay (38.5\% branching ratio) to \ce{^{64}Zn} \cite{Singh2007}. The emitted 190-keV $\beta^-$ particle makes this an  attractive short-range therapeutic radionuclide, which also has the possibility for simultaneous positron emission tomography (PET) imaging for real-time dose monitoring and verification. This makes \ce{^{64}Cu} particularly attractive for emerging radiation therapy protocols \cite{Lewis2003,NSACIsotopesSubcommittee2015,Bandari2014,mp500671j}. In addition, copper radiochemistry is well developed, and many existing ligands and carriers may be used for selective delivery of the radionuclide to different sites in patients. The second radionuclide studied, \ce{^{47}Sc} ($t_{1/2}$ = 3.35 d), undergoes $\beta^-$ decay to \ce{^{47}Ti}, emitting a high-intensity 159-keV gamma ray in the process \cite{Burrows2007}. This radionuclide is  attractive as an emerging diagnostic isotope, due to the similarity of the emitted gamma ray to that of the more well-established \ce{^{99m}Tc} \cite{Qaim2011,Qaim201731,Kolsky1998,mausner1995evaluation}. However, due to the short half-life of and dwindling supplies of \ce{^{99m}Tc}, \ce{^{47}Sc} stands poised as a potential solution to this shortage, due to its longer half-life and multiple production pathways without the need for highly enriched uranium. In addition, when paired with \ce{^{44}Sc}, \ce{^{47}Sc} forms a promising \enquote{theranostic} set for use in simultaneous therapeutic and diagnostic applications \cite{Muller2014,Deilami-nezhad2016}.

% \comment{cite theranostic papers, etc}

@@

% \comment{Regex to replace all hard figure references with LaTeX cross-references  [F,f]igure[^s*}]     }

@@

The approach used in both measurements was to irradiate foils of zinc or titanium, which were co-loaded with indium foils in order to determine their (n,p) cross sections relative to the well-established \ce{^{113}In}(n,n')\ce{^{113m}In} and \ce{^{115}In}(n,n')\ce{^{115m}In} neutron dosimetry standards \cite{Capote2012}.   \autoref{tab:foil_specs} records physical characteristics of each foil for the various irradiations. In each experiment, the co-loaded foils were irradiated for 3 hours at nominal operating conditions of 1.3 mA and 100 kV.  After irradiation, the foils were removed and placed in front of an appropriate High-Purity Germanium (HPGe) gamma-ray detector and the decay gamma-rays were recorded in variable-time bins.

@@

% \comment{Regex to replace table hard link with LaTeX cross-reference. - match w/ [T,t]able[^*}]   }

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
% \usepackage{graphicx}
\begin{table*}
\centering
\caption{Foil characteristics for each of the three (Zn/In)* experiments and the two (Ti/In)$^\dagger$  experiments. 
}
\label{tab:foil_specs}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}ccccccc@{}}
\toprule
Foils Used & Metal Purity          & Abundance (a/o)                                                           & \begin{tabular}[c]{@{}c@{}}Foil Density \\ (mg/cm$^2$)\end{tabular} & Thickness (mm)                                                                                                         & Diameter (cm)                                                                                                           & Mass (g, $\pm$ 5 mg)                                                                                    \\ \midrule
$^\text{nat}$In      & \textgreater 99.999\% & \begin{tabular}[c]{@{}c@{}}\ce{^{113}In} (4.29\%),\\ \ce{^{115}In} (95.71\%)\end{tabular} & 365.5                                                            & \begin{tabular}[c]{@{}c@{}}0.49 $\pm$ 0.02*, \\ 0.50 $\pm$ 0.03*, \\ 0.49 $\pm$ 0.03*, \\ 0.53 $\pm$ 0.06$^\dagger$, \\ 0.50 $\pm$ 0.02$^\dagger$\end{tabular} & \begin{tabular}[c]{@{}c@{}}9.75 $\pm$ 0.09*, \\ 9.98 $\pm$ 0.15*, \\ 9.96 $\pm$ 0.10*, \\ 10.01 $\pm$ 0.11$^\dagger$, \\ 10.00 $\pm$ 0.09$^\dagger$\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.248*, \\ 0.248*, \\ 0.241*, \\ 0.247$^\dagger$,\\  0.248$^\dagger$\ \end{tabular} \\ \hline
$^\text{nat}$Zn      & \textgreater 99.99\%  & \ce{^{64}Zn} (49.17\%)                                                            & 714.1                                                            & \begin{tabular}[c]{@{}c@{}}1.03 $\pm$ 0.01, \\ 1.03 $\pm$ 0.01, \\ 1.02 $\pm$ 0.01\end{tabular}                                    & \begin{tabular}[c]{@{}c@{}}9.93 $\pm$ 0.14, \\ 9.76 $\pm$ 0.17, \\ 9.89 $\pm$ 0.15\end{tabular}                                     & \begin{tabular}[c]{@{}c@{}}0.538, \\ 0.451, \\ 0.452\ \end{tabular}                           \\ \hline
$^\text{nat}$Ti      & 99.999\%              & \ce{^{47}Ti} (7.44\%)                                                             & 450.6                                                            & \begin{tabular}[c]{@{}c@{}}1.16 $\pm$ 0.02, \\ 1.15 $\pm$ 0.02\end{tabular}                                                         & \begin{tabular}[c]{@{}c@{}}9.93 $\pm$ 0.04, \\ 9.94 $\pm$ 0.03\end{tabular}                                                          & \begin{tabular}[c]{@{}c@{}}0.337, \\ 0.337\ \end{tabular}                                   \\ \bottomrule
\end{tabular}%
}
\end{table*}

% \comment{Need foil data from Mauricio's 2nd Ti run}

@@

\comment{Regex to replace equations with LaTeX formats, add cross-references.} % [E,e]quation[^*}]    }

@@

M. B. Chadwick,et al., Nucl. Data Sheets, vol. 107, no. 12, 2006.   (ENDF)

@@

% C. Müller et al., J. Nucl. Med., vol. 55, no. 10, Oct. 2014.

N. Otuka et al., Nucl. Data Sheets, vol. 120, 2014.   (EXFOR)

% L. Pietrelli et al., J. Radioanal. Nucl. Chem., vol. 157, no. 2, 1992.

% S. M. Qaim et al., IAEA Technical Reports Series No. 473, 2011.

@@

% D. Updegraff et al., ''Nuclear Medicine without Nuclear Reactors or Uranium Enrichment,'' 2013.
